---
title: "hw12.4"
output: word_document
date: "2023-11-27"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Simulation of Markov Process
a.
```{r}
library(expm)
matr=t(matrix(c(0.2,0.7,0.1,0.2,0.5,0.3,0.2,0.4,0.4),nrow=3,ncol=3))
print(matr)

```


1b.
```{r}
currentstate=1
update<- function(currentstate){
  new=sample(c(1,2,3),1,prob=matr[currentstate,])
  return(new)
}
update(currentstate)

```



2a. 
```{r}
values=eigen(t(matr))$values
vectors=eigen(t(matr))$vectors
print(values)
#first eigenvalue is equal to 1
#normalize corresponding eigenvector
stationary=vectors[,1]/sum(vectors[,1])
print(stationary)

```



2b. We can see that regardless of the initial probability distribution, the distributions converge to the stationary distribution.
```{r}
initial1=matrix(c(1,0,0),nrow=1)
initial2=matrix(c(0,1,0),nrow=1)
initial3=matrix(c(0,0,1),nrow=1)
i=50
norm1=c()
norm2=c()
norm3=c()
for(j in 1:i){
  new1=initial1%*%(matr%^%j)
  norm1[j]=sqrt(sum((new1-stationary)^2))
  new2=initial2%*%(matr%^%j)
  norm2[j]=sqrt(sum((new2-stationary)^2))
  new3=initial3%*%(matr%^%j)
  norm3[j]=sqrt(sum((new3-stationary)^2))
}
i=seq(1,50,1)
plot(i,norm1,col='red','l',log='y',xlab='iteration',ylab='Norm of Difference')
lines(i,norm2)
lines(i,norm3,col='blue')
legend('right',legend=c('starting from state1','starting from state2', 'starting from state3'),col=c('red','black','blue'),lty = c(1,1,1))


```
  
3a.
```{r}
set.seed(101)
trials=10000
counts1=c()
for(i in 1:trials){
state=1
counter=0
while(state!=3){
  state=update(state)
  counter=counter+1
}
counts1[i]=counter
}
counts2=c()
for(i in 1:trials){
state=2
counter=0
while(state!=3){
  state=update(state)
  counter=counter+1
}
counts2[i]=counter
}
mean1=mean(counts1)
mean2=mean(counts2)
hist(counts1)
abline(v=mean1)
hist(counts2,xlim=c(0,25))
abline(v=mean2)
#the histogram below shows the counts starting at state 2 in blue and starting at state1 in red.
hist(counts2,col=rgb(0,0,1,1/4))
hist(counts1,col=rgb(1,0,0,1/4),add=TRUE)
abline(v=mean1,col=rgb(1,0,0,1))
abline(v=mean2,col=rgb(0,0,1,1))

print(mean1);print(mean2)
```

3b. The means computed theoretically are quite close to the means found in our simulations. The theoretical mean starting from 1 is 4.615 compared to the simulation mean of4.745 and the theoretical mean for 2 is 3.846 compared to the simulation mean of 3.851.